A simulation model of DNA methylation data

### 2015-03-25 Rmarkdown/methsim.Rmd
Finished section on simulating Z.

\subsubsection{Simulating $\mathbf{{Z_{\tilde{h}}}}$}

@@

`methsim` assigns these transition probabilities based on $\beta_{i | r_{i}}$ and $LOR_{(i, i+1) | r_{i}, IPD(i, i+1)}$, the empirical distributions of $\beta_{i}$ and $LOR_{IPD(i, i + 1)}$ conditional on the region type of the $i^{th}$ locus, $r_{i}$, and the intra-pair distance of the the $i^{th}$ and $(i + 1)^{th}$ locus, $IPD(i, i+1)$, respectively. These empirical distributions are shown for the \emph{ADS} sample in Figures \ref{fig:ADS_beta_by_pm_boxplots} and \ref{fig:ADS_lor_by_pm_quantile_plots} (__TODO: Are these other examples necessary?__)

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/ADS_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{ADS} sample. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:ADS_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/ADS_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{ADS} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:ADS_lor_by_pm_quantile_plots}
\end{figure}

These same distributions for all the _Ziller_, _Lister_, _Seisenberger_ and _Ziller_ datasets are shown in Figures \ref{fig:EPISCOPE_beta_by_pm_boxplots}, \ref{fig:EPISCOPE_lora_by_pm_quantile_plots}, \ref{fig:Lister_beta_by_pm_boxplots}, \ref{fig:Lister_lora_by_pm_quantile_plots},
\ref{fig:Seisenberger_beta_by_pm_boxplots}, \ref{fig:Seisenberger_lor_by_pm_quantile_plots},
\ref{fig:Ziller_merged_beta_by_pm_boxplots} and \ref{fig:Ziller_merged_lor_by_pm_quantile_plots}.

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/EPISCOPE_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{EPISCOPE} dataset. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:EPISCOPE_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/EPISCOPE_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{EPISCOPE} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:EPISCOPE_lor_by_pm_quantile_plots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Lister_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{Lister} dataset. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Lister_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Lister_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{Lister} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Lister_lor_by_pm_quantile_plots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Seisenberger_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{Seisenberger} dataset. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Seisenberger_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Seisenberger_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{Seisenberger} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Seisenberger_lor_by_pm_quantile_plots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Ziller_merged_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{Ziller} dataset. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Ziller_merged_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Ziller_merged_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{i | r_{i}}$ and $LOR_{i | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{Ziller} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Ziller_merged_lor_by_pm_quantile_plots}
\end{figure}

@@

1. For each region, $r_{i}$, sample $B_{r_{i}}$ from $\beta_{i | r_{i}}$. This can be interpreted as $Pr(Z_{r_{i}} = 1)$, the marginal probability that a methylation locus in region $r_{i}$ is methylated, or, alternatively, as the average level of methylation for all methylation loci in the region.
2. For every pair of methylation loci, $(i, i + 1)$, $ 1 \leq i \leq (n - 1)$, sample $\psi_{(i, i + 1)}$ from $LOR_{(i, i+1) | r_{i}, IPD(i, i+1)}. This can be interpreted as the odds that the $(i + 1)^{th}$ methylation locus has the same methylation state as the $i^{th}$ methylation locus.

The second step requires further explanation. If there are an insufficient number of observations with the given $IPD$-region combination, then, by default, `methsim` samples $\psi_{(i, i + 1)}$ from a Gaussian distribution with mean equal to __TODO: What?__ and standard deviation equal to __TODO: What?__. __TODO: Explain how these means and standard deviations are estimated using loess.__ By default, if there are less than __TOOD: how many?__ $LOR$ with the given $IPD(i, i + 1), r_{i}$ combination then the Normal-approximation is used. The user may specify a different method for dealing with these rarer $IPD$-region combinations.

Now that we have $B_{r_{i}}$ and $\psi_{i, i + 1}$ for $1 \leq i \leq n - 1$ we can simulate realisations of $\mathbf{Z_{\tild{h}}}$. To begin the process (i.e., when $i = 1$) $Z_{\tilde{h}, 1}$, is simulated from the "marginal" distribution, which is Bernoulli($B_{r_{1}}$). Then, we loop over $i = 2, \ldots, n - 1$ to simulate the remainder of the process. At each step we compute the transition probability, $p_{i, i + 1} = Pr(Z_{\tilde{h}, i + 1} = 1 | Z_{\tilde{h}, i}) = \frac{Pr(Z_{tilde{h}, i}, Z_{tilde{h}, i + 1} = 1)}{Pr(Z_{tilde{h}})}$, from $B_{r_{i}}, $B_{r_{i + 1}}, \psi_{(i, i + 1)}$ and the observed (simulated) value at the previous iteration, $Z_{\tilde{h}, i}$. Then, $Z_{\tilde{h}, i + 1}$ is simulated from the Bernoulli($p_{i, i + 1}$) distribution.

To compute the transition probability, $p_{i, i + 1}$, `methsim` uses the iterative proportional fitting algorithm (IPF; __TODO: CITE IPD paper__) to construct the joint probability matrix, __TODO: Mathematically describe this matrix__. Then, the appropriate element of this matrix is divided by the appropriate marginal probability to obtain the transition probability. An example of this procedure is illustrative.

The IPD procedure is a general method for __TODO: blah blah__. In this case we wish to use the IPF procedure to construct the unique $2 \times 2$ array (the joint probability matrix) with specified margins (the marginal probabilities of each methylation state at each locus) and the specified cross-product ratio (the odds-ratio). __UP TO HERE__

\begin{equation}
% An example of IPF being used to form a transition matrix.
\end{equation}

To simulate $Z_{h, j}$ the algorithm proceeds as follows (in `R` pseudocode):

```{r, echo = TRUE, eval = FALSE}
n <- length(z)
u <- runif(n, 0, 1)
# Sample from region-specific beta-value distributions
b <- sample(BETA, n)
# Sample from region-specific LOR distributions
l <- sample(LOR, n - 1)

# Sample the first observation from the "stationary/marginal" distribution
if (u[1] < b[1]) {
  z[1] <- 0
} else {
  z[1] <- 1
}

# Sample the rest of the z
for (i in 2:n) {
  if (z[i] == 0) {
    # Compute transition probability
    # Compute z[i]

  } else {
    # Compute transition probability
    # Compute z[i]
  }
}
```

\subsubsection{Simulating a single sample}

@@

Given a set of transition matrices, $\Big \{ P_{i^{(h, j)}} \Big \}$, we can readily simulate a realisation of the methylation states along a haploid genome, $Z_{h, j}$. Naively, we might simulate a single sample by simply generating a realisation for each $h$ of $Z_{h, j}$. However, although there is cell-to-cell heterogeneity in the methylation patterns, there is still a degree of dependence for the $i^{th}$ methylation state across the $h$ haploid genomes in the $j^{th}$ sample. After all, the cells are from the same individual and are broadly of the same "type".

What this means is that we want some degree of dependence between the set of transition matrices for haploid methylome $h$, $\Big \{ P_{i^{(h, j)}} \Big \}$, and for those for haploid methylome $h'$, $\Big \{ P_{i^{(h', j)}} \Big \}$. This is asking quite a lot of our simple simulation model. Instead, `methsim` uses a heuristic procedure to achieve a similar result.

To begin, we assume that all $H$ haploid genomes in the sample have the same co-methylation structure. That is equivalent to defining $\Big \{ P_{i^{(h, j)}} \Big \} := \Big \{ P_{i^{(j)}} \Big \}$ for all $h$. __TODO: Comment on this assumption__

Naively, we might simulate a sample by simulating $H$ realisations of the Markov chain with transition matrices $\Big \{ P_{i^{(j)}} \Big \}$. However, this gives no control over the dependence between the methylation states at the $i^{th}$ locus over the $H$ haploid genomes in the $j^{th}$ sample. In fact, such a model will have the "correct" co-methylation structure but "incorrect" $\beta$-values. __TODO: Check this and demonstrate using `methsim`__

Instead, `methsim` simulates $R << H$ realisations of the Markov chain with transition probabilities given by $\Big \{ P_{i^{(j)}} \Big \}$. It then weights these realisations according to the observed heterogeneity for each region. __TODO: Describe the inuition of this weighting scheme as being that $w_{r}$ of the cells have identical methylation states for that region.

>>>>

\subsubsection{Simulating $\{ \mathbf{{Z_{\tilde{h}}}} \}_{\tilde{h} = 1}^{n_{\tilde{h}}}$}

@@

`methsim` assigns these transition probabilities based on $\beta_{i | r_{i}}$ and $LOR_{(i, i+1) | r_{i}, IPD(i, i+1)}$, the empirical distributions of $\beta_{i}$ and $LOR_{IPD(i, i + 1)}$ conditional on the region type of the $i^{th}$ locus, $r_{i}$, and the intra-pair distance of the the $i^{th}$ and $(i + 1)^{th}$ locus, $IPD(i, i+1)$, respectively. These empirical distributions are shown for the \emph{ADS} sample in Figures \ref{fig:ADS_beta_by_pm_boxplots} and \ref{fig:ADS_lor_by_pm_quantile_plots}.

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/ADS_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{ADS} sample. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:ADS_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/ADS_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{(i, i + 1) | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{ADS} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:ADS_lor_by_pm_quantile_plots}
\end{figure}

These same distributions for all the _Ziller_, _Lister_, _Seisenberger_ and _Ziller_ datasets are shown in Figures \ref{fig:EPISCOPE_beta_by_pm_boxplots}, \ref{fig:EPISCOPE_lor_by_pm_quantile_plots}, \ref{fig:Lister_beta_by_pm_boxplots}, \ref{fig:Lister_lor_by_pm_quantile_plots},
\ref{fig:Seisenberger_beta_by_pm_boxplots}, \ref{fig:Seisenberger_lor_by_pm_quantile_plots},
\ref{fig:Ziller_merged_beta_by_pm_boxplots} and \ref{fig:Ziller_merged_lor_by_pm_quantile_plots} (__TODO: Are these other examples necessary?__).

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/EPISCOPE_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{EPISCOPE} dataset. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:EPISCOPE_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/EPISCOPE_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{(i, i + 1) | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{EPISCOPE} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:EPISCOPE_lor_by_pm_quantile_plots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Lister_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{Lister} dataset. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Lister_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Lister_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{(i, i + 1) | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{Lister} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Lister_lor_by_pm_quantile_plots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Seisenberger_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{Seisenberger} dataset. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Seisenberger_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Seisenberger_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{(i, i + 1) | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{Seisenberger} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Seisenberger_lor_by_pm_quantile_plots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Ziller_merged_beta_by_pm_boxplots.pdf}
\caption{$\beta_{i | r_{i}}$, the distribution of $\beta$-values for CpGs in each region type, for the \emph{Ziller} dataset. Only CpGs with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Ziller_merged_beta_by_pm_boxplots}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/Ziller_merged_lor_by_pm_quantile_plots.pdf}
\caption{$LOR_{(i, i + 1) | r_{i}}$, the distribution of $\log_{2}$ odds-ratio estimates of within-fragment co-methylation for CpGs in each region type, for the \emph{Ziller} sample. Only 2-tuples with at least $10\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:Ziller_merged_lor_by_pm_quantile_plots}
\end{figure}

@@

1. For each region, $r_{i}$, sample $B_{r_{i}}$ from $\beta_{i | r_{i}}$. This can be interpreted as $Pr(Z_{r_{i}} = 1)$, the marginal probability that a methylation locus in region $r_{i}$ is methylated, or, alternatively, as the average level of methylation for all methylation loci in the region[^extreme_B].
2. For every pair of methylation loci, $(i, i + 1)$, $1 \leq i \leq (n - 1)$, sample $\Psi_{(i, i + 1)}$ from $LOR_{(i, i+1) | r_{i}, IPD(i, i+1)}$. This can be interpreted as the odds that the $(i + 1)^{th}$ methylation locus has the same methylation state as the $i^{th}$ methylation locus.

[^extreme_B]: If $B_{r_{i}} = 0$ (resp. $B_{r_{i}} = 1$), `methsim` adds (resp. subtracts) a small value $\epsilon$. The default value is $\epsilon = 0.01$. The boundary values of $0$ and $1$ are common in $\beta_{i | r_{i}}$ and, were these retained, would result in methylomes with large regions that are uniformly unmethylated (resp. methylated).

The second step requires further explanation. If there are an insufficient number of observations with the given $IPD$-region combination, then, by default, `methsim` samples $\Psi_{(i, i + 1)}$ from a Gaussian distribution with mean equal to __TODO: What?__ and standard deviation equal to __TODO: What?__. __TODO: Explain how these means and standard deviations are estimated using loess.__ By default, if there are less than __TOOD: how many?__ $LOR$ with the given $IPD(i, i + 1), r_{i}$ combination then the Normal-approximation is used. The user may specify a different method for dealing with these rarer $IPD$-region combinations.

Now that we have $B_{r_{i}}$ and $\Psi_{i, i + 1}$ for $1 \leq i \leq n - 1$ we can simulate realisations of $\mathbf{Z_{\tilde{h}}}$. To begin the process (i.e., when $i = 1$) $Z_{\tilde{h}, 1}$, is simulated from the "marginal" distribution, which is Bernoulli($B_{r_{1}}$). Then, we loop over $i = 2, \ldots, n - 1$ to simulate the remainder of the process. At each step we compute the transition probability, $p_{(i + 1) | i} = Pr(Z_{\tilde{h}, i + 1} = 1 | Z_{\tilde{h}, i}) = \frac{Pr(Z_{tilde{h}, i}, Z_{tilde{h}, i + 1} = 1)}{Pr(Z_{tilde{h}})}$, from $B_{r_{i}}, B_{r_{i + 1}}, \Psi_{(i, i + 1)}$ and the observed (simulated) value at the previous iteration, $Z_{\tilde{h}, i}$. Then, $Z_{\tilde{h}, i + 1}$ is simulated from the Bernoulli($p_{(i + 1) | i}$) distribution.

In order to compute the transition probability, $p_{(i + 1) | i}$, `methsim` first uses the iterative proportional fitting algorithm \citet[see]{Speed:2005tg}[for an overview of the algorithm] to construct the joint probability matrix,

\begin{equation*}
P = \bordermatrix{~ & Pr(Z_{i+1} = 0) = 1 - B_{r_{i+1}} & Pr(Z_{i+1} = 1) = B_{r_{i+1}} \cr
Pr(Z_{i} = 0) = 1 - B_{r_{i}} & Pr(Z_{i} = 0, Z_{i + 1} = 0) & Pr(Z_{i} = 0, Z_{i + 1} = 1) \cr
Pr(Z_{i} = 1) = B_{r_{i}} & Pr(Z_{i} = 1, Z_{i + 1} = 0) & Pr(Z_{i} = 1, Z_{i + 1} = 1) \cr}
\end{equation*}

Iterative proportional fitting is a general method "for constructing tables of numbers satisfying certain constraints" \citep{Speed:2005tg}. `methsim` uses iterative proportional fitting to construct the _unique_ $2 \times 2$ array (the joint probability matrix) with specified margins (the marginal probabilities of each methylation state at each locus, $B_{r_{i}} = Pr(Z_{i} = 1), B_{r_{i + 1}} = Pr(Z_{i + 1} = 1)$) and the specified cross-product ratio (the odds-ratio, $\psi_{i, i + 1} = \log_{2}(\Psi_{i, i + 1})$).

Once $P$ is formed, the appropriate element of it is divided by the appropriate marginal probability to obtain the transition probability. An example of this procedure is illustrative.

Let $\Psi_{i, i + 1} = 2$, and therefore $\psi = 2^{2} = 4$, and $B_{r_{i}} = B_{r_{i + 1}} = 0.7$. To begin the iterative proportional fitting algorithm, form the matrix $P^{(0)} = \Bigl(\begin{smallmatrix}
\psi & 1 \\
   1 & 1
\end{smallmatrix} \Bigr).$ This matrix has the desired cross-ratio, $\psi$, but not the desired row and column margins. At each iteration of the algorithm, iterative proportional fitting adjusts the rows and columns such that the cross-ratio remains $\psi$ while the row and column margins converge towards the desired values. The algorithm continues in this manner, forming a series of $2 \times 2$ tables that converge pointwise (and uniquely) to a $2 \times 2$ table $P^{(*)} \equiv P$. For our example, $P^{(*)} = \Bigl(\begin{smallmatrix}
0.122 & 0.178 \\
0.178 & 0.522
\end{smallmatrix} \Bigr)$, to 3 decimal places of precision. It is easily verified that $P^{(*)}$ has the desired cross-ratio and marginal sums.

In `methsim`, the routine to simulate $\{ \mathbf{{Z_{\tilde{h}}}} \}_{\tilde{h} = 1}^{n_{\tilde{h}}}$ is available as the `simulate,SimulateMethylomeParam-method`. It is mostly written in C++ using the wonderful Rcpp package \citep{Eddelbuettel:2011to, Eddelbuettel:2013if}, which greatly speeds up the running time.

\subsubsection{Simulating $\mathbf{W}$}

@@

For each region we need to assign the vector of weights, $\mathbf{W_{r_{i}, \cdot}}$. To do so, `methsim` uses a heuristic procedure. For higher-order m-tuples, $m = 3$ or $4$ is recommended, `methsim` computes the relative frequency of the $2^{m}$ possible methylation patterns at each m-tuple with sufficient sequencing coverage. `methsim` then samples from this distribution, stratified by region-type; call these empirical distributions

__TODO__: Really computing (and plotting) $w_{i | r_{(i, \ldots, i + m - 1)}}, \ldots, w_{i + m - 1 | r_{(i, \ldots, i + m - 1)}}$ not $w_{(i, \ldots, i + m - 1) | r_{(i, \ldots, i + m - 1)}}$. These empirical distributions are shown for the \emph{ADS} sample in Figure \ref{fig:ADS_pattern_freqs_by_pm_boxplots}. From this we see that __TODO__.

__TODO: Update figures to use $w_{1}, \ldots, w_{8}$ rather than $h1, \ldots h8$.__

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{../figures/ADS_pattern_freqs_by_pm_boxplots.pdf}
\caption{$w_{(i, \ldots, i + m - 1) | r_{(i, \ldots, i + m - 1)}}$, the distribution of the relative frequencies for all CpG 3-tuples in each region type, for the \emph{ADS} sample. Only CpGs with at least $5\times$ sequencing coverage are used. UMR = unmethylated region; LMR = lowly methylated region; PMR = partially methylated region; other = any other region.}
\label{fig:ADS_beta_by_pm_boxplots}
\end{figure}

__TODO: Limitations of this procedure__
